State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-28 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex

%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx,epsfig}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
\usepackage[tight,scriptsize]{subfigure}
\usepackage{pstool}
%\renewcommand{\subcapsize}{\small}
% \usepackage{anysize}							%<-use this for checking equation length
% \marginsize{1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
% \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
% \usepackage{amsmath,amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\wtf}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\omg}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} \begin{document}

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}

%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'. 
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right). 
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{equation}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}\label{eq:FieldCovarianceFunction}
	cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ 
	\begin{array}{*{20}
		{c}} \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
		0 & \mathrm{otherwise}. \\
	\end{array}
	\right. 
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

	\label{reduced continuous model} \boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T_s\int_\Omega { f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right)^\top d\mathbf{r}'}\boldsymbol{\theta} - \xi\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right). 

>>>>

%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx,epsfig}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
\usepackage[tight,scriptsize]{subfigure}
% \usepackage{pstool}
\renewcommand{\subcapsize}{\small}
% \usepackage{anysize}							%<-use this for checking equation length
% \marginsize{1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
% \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
\usepackage{amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\wtf}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\omg}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} \begin{document}

% \begin{document}

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{eqnarray}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =  \int_{-\infty}^t h\left(t - t'\right) \\
	\times\left(\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'. 
\end{eqnarray}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{eqnarray}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \\
	\int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right). 
\end{eqnarray}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{eqnarray}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} \\
	+ T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right), 
\end{eqnarray}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}\label{eq:FieldCovarianceFunction}
	cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ 
	\begin{array}{*{20}
		{c}} \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
		0 & \mathrm{otherwise}. \\
	\end{array}
	\right. 
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

	\label{reduced continuous model} \boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T_s\int_\Omega { f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right)^{\top}d\mathbf{r}'}\boldsymbol{\theta} \\
	- \xi\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right). 
